---
title: "Hospitalisations paper scripts"
author: "Jan Savinc"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---


# Overview

Here I process data from the safe haven into neat tables and graphs for the hospitalisations paper


# Preliminaries

```{r}
library(tidyverse)
library(patchwork)

source("./FUNCTIONS.R")  # load common functions
```


# Load data

This may be slow as I'm reading from the X:/ drive

```{r}
tbl_paths <- 
  tribble(
    ~model, ~path,
    "hosp_poisson", "X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_hosp_poisson.csv",
    "hosp_negbin", "X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_hosp_negbin.csv",
    "hosp_zinb", "X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_hosp_zinb.csv",
    "los_poisson", "X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_los_poisson.csv",
    "los_negbin", "X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_los_negbin.csv",
    "los_zinb", "X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_los_zinb.csv",
  )

model_coefficients <- 
  pmap(
  .l = tbl_paths,
  .f = ~read_csv(file = .y)
) %>%
  set_names(nm = tbl_paths$model)

model_coefficients_rse <- read_csv("X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_coefficients_rse.csv")
model_coefficients_other <- read_csv("X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_coefficients_other.csv")
model_coefficients_ols <- read_csv("X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_coefficients_ols.csv")

model_ames <- read_csv("X:\\R2090\\2021-0312 Deaths at home\\safe_haven_exports\\modelling_hospitalisations\\tbl_models_ames.csv")
```

# Compare coefficients

For hospitalisations, the coefficients are very similar

```{r}
model_coefficients %>%
  list_rbind(names_to = "model") %>% 
  filter(str_detect(model, "hosp\\_")) %>%
  filter(is.na(Component) | Component == "conditional") %>%
  mutate(
    IRR = coalesce(IRR, exp_coef),
    IRR_CI_low = coalesce(IRR_CI_low, exp_CI_low),
    IRR_CI_high = coalesce(IRR_CI_high, exp_CI_high),
    intercept = variable=="(Intercept)"
  ) %>%
  ggplot(
    aes(x = IRR, y = paste0(variable,": ",level), xmin = IRR_CI_low, xmax = IRR_CI_high, shape = model)
  ) +
  geom_vline(xintercept = 1, linetype = "dashed", colour = "grey75") +
  geom_point(size = rel(2), position = position_dodge(width = 1)) +
  geom_linerange(position = position_dodge(width = 1)) +
  facet_wrap(vars(intercept), scales = "free", nrow = 2) +
  NULL
```


# Compare AMEs

This is the main way I'm comparing between the different models since they adopt different methods: to compare between the zero-inflated negative binomial and all the others a common basis is needed, because the zero-inflated model produces 2 sets of coefficients, and they are not easily comparable

## Days between discharge & death

```{r}

# TODO: the CIs on the Poisson look too small, plus it looks like the RSE versions of the Poisson AMEs aren't there?!
# TODO: check the SH to see if something went wrong when computing these?
# TODO: maybe change HC0 to HC3, or just set it to "sandwich" or whatever the default is? I just want a HC estimator of variance, don't really care which

model_ames %>%
  filter(str_detect(model, "discharge")) %>%
  ggplot(
    aes(x = estimate, y = paste0(term,": ",contrast), shape = model, xmin = conf.low, xmax = conf.high)
  ) +
  geom_point(size = rel(2), position = position_dodge(width = 1)) +
  geom_linerange(position = position_dodge(width = 1))
```

# Pretty coefficient table, using Poisson + RSE

```{r}
# TODO: change pandemic variable to have pre-pandemic as the reference...
(tbl_model_coefficients_poisson_rse <-
  model_coefficients_rse %>%
  mutate(
    variable = replace_with_long_variable_name(variable),
    p = scales::label_pvalue()(p.value),
    ci = paste0(signif(conf.low, 3), "-", signif(conf.high, 3))
    ) %>%
  select(-c(statistic, std.error, conf.low, conf.high, p.value)) %>%
  mutate(model = str_remove(model, "\\_poisson\\_rse")) %>%
  pivot_wider(names_from = model, values_from = c(estimate, ci, p), names_vary = "slowest"))

view(tbl_model_coefficients_poisson_rse, title = "IRRs")
```

