---
title: "Hospitalisations paper scripts, using models with time term"
author: "Jan Savinc"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---
  

# Overview

This is a new script processing the outputs from the hospitalisations modelling analysis, based on the outputs from October 2026 - this includes models with an added time term, models with all places of death included, and an interaction term between the pandemic period and cause of death.



# Preliminaries

```{r}
library(tidyverse)
library(patchwork)
library(broom)  # for tidy()
library(tictoc)  # for timing
library(janitor)  # for some nice cleaning functions
# library(flowchart)  # for plotting a flowchart; not usable for me - it's good if you have raw data!
library(ggflowchart)  # for plotting a flowchart

source("./FUNCTIONS.R")  # load common functions

dir_output <- "C:/Users/40011625/OneDrive - Edinburgh Napier University/SCADR/covid_mortality_home/dissemination/drafts//hospitalisations paper/"
```

## Functions & constants

```{r}
# specify the order of place of death
levels_place_of_death <- c("Hospital", "Home", "Care home", "All")

levels_cod_nrs <- c("Cancer", "Circulatory", "Dementia & Alzheimer's", "Respiratory", 
"Other", "COVID-19", "All causes")

code_place_of_death_consistently <- function(data_tbl) {
  data_tbl %>%
    mutate(
      ## shorten to "Home"
      cat_place_of_death = if_else(cat_place_of_death == "Home & non-institution", true = "Home", false = cat_place_of_death),
      ## Shorten to "Care home"
      cat_place_of_death = if_else(cat_place_of_death == "Care home & other", true = "Care home", false = cat_place_of_death),
      cat_place_of_death = factor(cat_place_of_death, levels = levels_place_of_death)
    )
}

## helper function to merge all pre-pandemic years into one
merge_years_n <- function(data_tbl) {
  data_tbl %>%
    mutate(
      val_cohort_year = fct_collapse(.f = val_cohort_year, "2020-21" = c("2020-21"), other_level = "2015-20") %>%
        fct_relevel("2015-20")
    ) %>%
    group_by(val_cohort_year, .add = TRUE) %>%
    summarise(n = sum(n), .groups = "drop")
}
merge_years_m_sd <- function(data_tbl) {
  data_tbl %>%
    mutate(
      val_cohort_year = fct_collapse(.f = val_cohort_year, "2020-21" = c("2020-21"), other_level = "2015-20") %>%
        fct_relevel("2015-20")
    ) %>%
    group_by(val_cohort_year, .add = TRUE) %>%
    summarise(mean = weighted.mean(mean, n_deaths),
              sd = sd_pooled(sd, n_deaths),
              n_deaths = sum(n_deaths),  # retain the number of individuals in each merged category - for hypothesis testing later!
              .groups = "drop")
}
```


# Load data & wrangle

This may be slow as I'm reading from the X:/ drive.

## Regression data & AMEs

```{r}
model_coefficients_rse_univariate <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_univariate_regressions_poisson_rse.csv")

model_coefficients_rse <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_models_coefficients_rse.csv")

model_coefficients_rse_with_interaction <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_models_coefficients_with_cod_interaction_rse.csv")

model_ames_rse <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_models_ames_rse.csv")

model_ames_rse_with_interaction <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_models_ames_rse_with_cod_interaction.csv")

model_collinearity <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_models_collinearity.csv")

model_metrics <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_models_fit_metrics.csv")

model_lrtests <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_models_rse_lrtests.csv")

model_sequential_check <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/modelling_hospitalisations/tbl_models_sequential_adjustment_rse.csv")
```


# Nicely formatted coefficient tables

## Univariate

```{r}

```

## Poisson + RSE

```{r}
# TODO: refresh the code below to work with the column names!
# TODO: update naming so I get a sensible interaction name
(tbl_model_coefficients_poisson_rse <-
  model_coefficients_rse %>%
  mutate(
    variable = replace_with_long_variable_name(variable),
    p_pretty = scales::label_pvalue()(p.value),
    stars = gtools::stars.pval(p.value),
    IRR = paste0(signif(IRR, 3), stars),
    ci = format_ci(IRR_CI_low, IRR_CI_high),
    ) %>%
  select(c(model, variable, level, IRR, ci)) %>%
  mutate(model = str_remove(model, "\\_poisson\\_rse")) %>%
  pivot_wider(names_from = model, values_from = c(IRR, ci), names_vary = "slowest"))

tbl_model_coefficients_poisson_rse %>%
  mutate(variable = if_else(condition = duplicated(variable), true = " ", false = as.character(variable))) %>%
  write_csv(., file = file.path(dir_output, "tbl_model_coefficients_poisson_rse.csv"))
```


