---
title: "Social Care analysis script for IPDLN 2026 poster"
author: "Jan Savinc"
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

# Preliminaries

Just a point to note that I use the terms social care and home care interchangeable throughout these analyses. I requested data from SOURCE on client characteristics, demographics (ethnicity), home care, and IORN. When I requested this, I thought this would result in overlapping but distinct populations - i.e. the demographics and client characteristics would include the largest number of people (also those in care homes and not receiving home care, for example), and it would get smaller for something like IORN.

In practice, the sample I've received is mostly home care users - of the people who had any record in the SOURCE extract, 98.5% had a home care record (the others only having a demographics record, demographics and client record, or demographics, client, and IORN). In other words, in this project `social care recipient` and `home care recipient` are practically interchangeable terms.

## Packages

```{r}
library(tidyverse)
library(patchwork)
library(broom)  # for tidy()
library(tictoc)  # for timing
library(janitor)  # for some nice cleaning functions

source("./FUNCTIONS.R")  # load utility functions
```

## Functions & constants

```{r}
# where to save outputs
dir_output <- "C:/Users/40011625/OneDrive - Edinburgh Napier University/SCADR/covid_mortality_home/dissemination/drafts/IPDLN 2026/"

# specify the order of place of death
levels_place_of_death <- c("Hospital", "Home", "Care home", "All")

code_place_of_death_consistently <- function(data_tbl) {
  data_tbl %>%
    mutate(
      ## shorten to "Home"
      cat_place_of_death = if_else(cat_place_of_death == "Home & non-institution", true = "Home", false = cat_place_of_death),
      ## Shorten to "Care home"
      cat_place_of_death = if_else(cat_place_of_death == "Care home & other", true = "Care home", false = cat_place_of_death),
      cat_place_of_death = factor(cat_place_of_death, levels = levels_place_of_death)
    )
}

# set plot font size
theme_update(text = element_text(size = 20))
```

## Load data

```{r}
# sense-check population
source_crosstab <- read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/crosstab_individuals_by_type_of_source_record.csv")

source_crosstab %>% group_by(bin_homecare) %>% summarise(n=sum(n), percent = scales::label_percent(accuracy=0.1)(sum(prop)))

demographics_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_characteristics_of_soc_care_recipients_vs_nonrecipients.csv") %>%
  mutate(val_cohort_year = factor(val_cohort_year, levels = order_cohort_year))

comorbidities_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_freq_comorbidities_recipients_vs_nonrecipients.csv") %>%
  mutate(val_cohort_year = factor(val_cohort_year, levels = order_cohort_year))

# this file contains descriptive stats of age, count of comorbid conditions, and the comorbidity index - I'll load it once, split it up into three, and delete the original file
age_and_comorb_index_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_age_comoorb_index_recipients_vs_nonrecipients.csv")

age_descriptives_by_recipient <- age_and_comorb_index_by_recipient %>% select(val_cohort_year, cat_soc_care_recipient, matches("^amt\\_age"))
comorb_count_descriptives_by_recipient <- age_and_comorb_index_by_recipient %>% select(val_cohort_year, cat_soc_care_recipient, matches("^val\\_comorb\\_count"))
comorb_index_descriptives_by_recipient <- age_and_comorb_index_by_recipient %>% select(val_cohort_year, cat_soc_care_recipient, matches("^val\\_elixhauser_index"))

rm(age_and_comorb_index_by_recipient)  # delete, no longer needed

cod_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_cod_recipients_vs_nonrecipients.csv") %>%
  mutate(
    val_cohort_year = factor(val_cohort_year, levels = order_cohort_year),
    cat_cod_nrs = factor(cat_cod_nrs, levels = unique(cat_cod_nrs))
    )

# Note: there is an error in the code producing this output - the "Covid-19 involved" category is the same as the "Covid underlying cause" category - this should have been different. I'll remove the "involved" category for now
covid_by_recipient_2020_21 <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_prop_soc_care_amongst_covid_deaths.csv") %>%
  filter(definition!="Covid-19 involved")

los_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_los_by_soc_care_receipt_and_time.csv") %>%
  mutate(val_cohort_year = factor(val_cohort_year, levels = order_cohort_year))
spells_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_n_spells_by_soc_care_receipt_and_time.csv") %>%
  mutate(val_cohort_year = factor(val_cohort_year, levels = order_cohort_year))
percent_admitted_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_prop_with_admissions_by_sc_receipit_and_time.csv") %>%
  mutate(val_cohort_year = factor(val_cohort_year, levels = order_cohort_year))
icd_chapters_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_icd_chapters_admissions_by_soc_care_receipt.csv") %>%
  mutate(
    val_cohort_year = factor(val_cohort_year, levels = order_cohort_year),
    chapter_icd10 = factor(chapter_icd10, levels = unique(chapter_icd10))
    )
```

## Wrangle data

### Consistent place of death name in demographics

```{r}
demographics_by_recipient <-
  demographics_by_recipient %>%
  mutate(
    level = case_match(level,
                       "Home & non-institution" ~ "Home",
                       "Care home & other" ~ "Care home",
                       .default = level)
  )
```


### Compute year-on-year changes in demographics

Merge the previous years' proportion and compute the difference. Start by deriving the previous year!

```{r}
demographics_by_recipient <-
  demographics_by_recipient %>%
  mutate(prev_year = order_cohort_year[as.integer(val_cohort_year)-1L]) %>%
  left_join(demographics_by_recipient %>% select(val_cohort_year:level,prop_previous = prop), by = join_by(prev_year==val_cohort_year, cat_soc_care_recipient, variable, level)) %>%
  mutate(
    prop_previous = if_else(val_cohort_year=="All", NA_real_, prop_previous),  # set to NA for the All years to avoid nonsensical results
    prop_change = prop - prop_previous
    ) %>%
  select(-c(prev_year,prop_previous))
```


### Compute year-on-year changes in comorbidities

Merge the previous years' proportion and compute the difference. Start by deriving the previous year!

```{r}
comorbidities_by_recipient <-
  comorbidities_by_recipient %>%
  mutate(prev_year = order_cohort_year[as.integer(val_cohort_year)-1L]) %>%
  left_join(comorbidities_by_recipient %>% select(val_cohort_year:cat_comorbidity,perc_previous = perc), by = join_by(prev_year==val_cohort_year, cat_soc_care_recipient, cat_comorbidity)) %>%
  mutate(
    perc_previous = if_else(val_cohort_year=="All", NA_real_, perc_previous),  # set to NA for the All years to avoid nonsensical results
    perc_change = perc - perc_previous
    ) %>%
  select(-c(prev_year,perc_previous))
```



### Top 12 most frequent comorbidities

I derive these from the "All" cohort year and "Both" social care recipient categories...

```{r}
top_12_comorb <-
  comorbidities_by_recipient %>%
  filter(val_cohort_year=="All") %>%
  count(cat_comorbidity, wt = n) %>%
  slice_max(order_by = n, n = 12) %>% pull(cat_comorbidity)
```


### Compute year-on-year changes in causes of death

Merge the previous years' proportion and compute the difference. Start by deriving the previous year!

```{r}
cod_by_recipient <-
  cod_by_recipient %>%
  mutate(prev_year = order_cohort_year[as.integer(val_cohort_year)-1L]) %>%
  left_join(cod_by_recipient %>% select(val_cohort_year:cat_cod_nrs,perc_previous = perc), by = join_by(prev_year==val_cohort_year, cat_soc_care_recipient, cat_cod_nrs)) %>%
  mutate(
    perc_previous = if_else(val_cohort_year=="All", NA_real_, perc_previous),  # set to NA for the All years to avoid nonsensical results
    perc_change = perc - perc_previous,
    perc_change = if_else(cat_cod_nrs=="COVID-19" & val_cohort_year=="2020-21", perc, perc_change)  # for Covid, percent change should equal the latest percentage
    ) %>%
  select(-c(prev_year,perc_previous))
```




# Outputs

## Miscellaneous

### Proportion of soc care recipients amongst covid deaths

```{r}
covid_by_recipient_2020_21 %>% knitr::kable(caption = "Proportion of social care (home care) recipients amongst people who died with Covid-19 listed as the underlying cause of death.")
```

### Mean count of comorbid conditions

```{r}
comorb_count_descriptives_by_recipient %>% knitr::kable(caption = "Descriptive statistics of the number of comorbid conditions (out of 31 included in the Elixhauser index) per person, over time, by social care (home care) recipient status.")
```

### Mean age of soc care recipients

```{r}
age_descriptives_by_recipient %>% knitr::kable(caption = "Descriptive statistics of the age at death, over time, by social care (home care) recipient status.")
```


## Demographics over time

- in 2020/21 age distribution has changed very little
- in 2020/21 sex shifted towards higher proportion in males in both recipients and non-recipients
- in 2020/21 SIMD has seen a decrease in 2nd quintile and increase in 1st quintile, while 3-5 remained similar
- not much changed on marital status
- ethnicity hard to show anything due to high degree of missingness...


```{r}
subplot_data_extract <- function(which_var) {
  demographics_by_recipient %>% 
      filter(variable == which_var) %>%
      filter(val_cohort_year == "2020-21") %>%
      filter(cat_soc_care_recipient!="Both") %>%
      mutate(level = factor(level, levels=unique(level)))  # set order of factor levels as it was exported
}

subplot_ggplot <- function(data, y = "prop") {
  ggplot(
    data = data,
    aes(x = level, y = .data[[y]], fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
    ) +
    geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(x=NULL,y=NULL) +
    theme(plot.title = element_text(hjust = 0.5)) +
    scale_fill_manual(values = order_colours_adr_s[1:2])
}

(subplot_age <-
  wrap_plots(
    subplot_data_extract("cat_age_decades") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_age_decades") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Age groups", theme = theme(plot.title = element_text(hjust=0.5)))

(subplot_sex <-
  wrap_plots(
    subplot_data_extract("cat_sex") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_sex") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Sex", theme = theme(plot.title = element_text(hjust=0.5)))

(subplot_marstat <-
  wrap_plots(
    subplot_data_extract("cat_marital_status_condensed") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_marital_status_condensed") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Marital status", theme = theme(plot.title = element_text(hjust=0.5)))

(subplot_simd <-
  wrap_plots(
    subplot_data_extract("cat_simd_quintile") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_simd_quintile") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Deprivation (SIMD quintile)", theme = theme(plot.title = element_text(hjust=0.5)))

(subplot_ur2 <-
  wrap_plots(
    subplot_data_extract("cat_ur2") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_ur2") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Urban-rural location", theme = theme(plot.title = element_text(hjust=0.5)))

(subplot_pod <-
  wrap_plots(
    subplot_data_extract("cat_place_of_death") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_place_of_death") %>%
      subplot_ggplot(y = "prop_change") +
      # scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Place of death", theme = theme(plot.title = element_text(hjust=0.5)))

# TODO: I lose the plot_annotation title for the subplots in this step...
wrap_plots(
  list(
    subplot_age, subplot_sex,
    subplot_pod, subplot_marstat,
    subplot_simd, subplot_ur2
  ), ncol = 2, nrow = 3
)
```

## Comorbid conditions over time

Only a few comorbidities that were more common in social care non-recipients... the individuals affected likely die before reaching social care age.

- solid tumour without metastasis
- metastatic cancer
- alcohol abuse
- liver disease
- lymphoma
- psychoses
- drug abuse
- aids/hiv

```{r}
# prints the comorbidities more common in non-recipients
comorbidities_by_recipient %>% select(-c(n, perc_change)) %>% pivot_wider(names_from = cat_soc_care_recipient, values_from = perc) %>% filter(Recipient < `Non-recipient`) %>% arrange(desc(Recipient)) %>% print_all

# TODO: decide how to present this... only the most frequent comorbs, but also the ones with the biggest difference between recipients and non-recipients?
# TODO: changes in pandemic weren't massive, not sure how to present those...

(subplot_comorbidities <-
  wrap_plots(
    comorbidities_by_recipient %>% 
      filter(val_cohort_year == "2020-21") %>%
      filter(cat_comorbidity %in% top_12_comorb) %>%
      mutate(cat_comorbidity = factor(cat_comorbidity, levels=unique(cat_comorbidity))) %>%  # set order of factor levels as it was exported
      ggplot(aes(x=cat_comorbidity,y=perc,fill=cat_soc_care_recipient,group=cat_soc_care_recipient)) +
      geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
      scale_y_continuous(labels = scales::label_percent()) +
      theme(plot.title = element_text(hjust = 0.5)) +
      labs(x=NULL,y=NULL,title="2020-21") +
      scale_fill_manual(values = order_colours_adr_s[1:2])
    ,
    comorbidities_by_recipient %>% 
      filter(val_cohort_year == "2020-21") %>%
      filter(cat_comorbidity %in% top_12_comorb) %>%
      mutate(cat_comorbidity = factor(cat_comorbidity, levels=unique(cat_comorbidity))) %>%  # set order of factor levels as it was exported
      ggplot(aes(x=cat_comorbidity,y=perc_change,fill=cat_soc_care_recipient,group=cat_soc_care_recipient)) +
      geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
      scale_y_continuous(labels = scales::label_percent()) +
      theme(plot.title = element_text(hjust = 0.5)) +
      labs(x=NULL,y=NULL,title="Percentage point change\nfrom 2019-20") +
      scale_fill_manual(values = order_colours_adr_s[1:2])
    ,
    guides = "collect"
    ) &
    plot_annotation(title = "Comorbid conditions", theme = theme(plot.title = element_text(hjust=0.5)))
)

(subplot_comorb_index <-
    comorb_index_descriptives_by_recipient %>%
    filter(val_cohort_year!="All") %>%
    filter(cat_soc_care_recipient!="Both") %>%
    ggplot(aes(x=val_cohort_year,y=val_elixhauser_index_vanwalraven_m,ymin=val_elixhauser_index_vanwalraven_ci_lo,ymax=val_elixhauser_index_vanwalraven_ci_hi, group=cat_soc_care_recipient, shape=cat_soc_care_recipient, colour=cat_soc_care_recipient)) +
  geom_point(size=rel(2)) +
  geom_line() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_manual(values = order_colours_adr_s[1:2]) +
  labs(x=NULL,y=NULL,title="Mean Elixhauser Comorbidity index")
)

(subplot_comorb_count <-
    comorb_count_descriptives_by_recipient %>%
    filter(val_cohort_year!="All") %>%
    filter(cat_soc_care_recipient!="Both") %>%
    ggplot(aes(x=val_cohort_year,y=val_comorb_count_m,ymin=val_comorb_count_ci_lo,ymax=val_comorb_count_ci_hi, group=cat_soc_care_recipient, shape=cat_soc_care_recipient, colour=cat_soc_care_recipient)) +
  geom_point(size=rel(2)) +
  geom_line() +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_manual(values = order_colours_adr_s[1:2]) +
  labs(x=NULL,y=NULL,title="Mean count of comorbid conditions")
)
```

## Causes of death over time

```{r}
(subplot_cod <-
  wrap_plots(
    cod_by_recipient %>% 
      filter(val_cohort_year == "2020-21") %>%
      ggplot(aes(x=cat_cod_nrs,y=perc,fill=cat_soc_care_recipient,group=cat_soc_care_recipient)) +
      geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
      scale_y_continuous(labels = scales::label_percent()) +
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_fill_manual(values = order_colours_adr_s[1:2]) +
      labs(x=NULL,y=NULL,title="2020-21")
    ,
    cod_by_recipient %>% 
      filter(val_cohort_year == "2020-21") %>%
      ggplot(aes(x=cat_cod_nrs,y=perc_change,fill=cat_soc_care_recipient,group=cat_soc_care_recipient)) +
      geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
      scale_y_continuous(labels = scales::label_percent()) +
      theme(plot.title = element_text(hjust = 0.5)) +
      scale_fill_manual(values = order_colours_adr_s[1:2]) +
      labs(x=NULL,y=NULL,title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    ) &
    plot_annotation(title = "Causes of death", theme = theme(plot.title = element_text(hjust=0.5)))
)
```


## Hospital use over time

### Number of hospital spells

```{r}
(subplot_hospital_spells <-
    spells_by_recipient %>%
    filter(cat_time==365) %>%
    filter(val_cohort_year!="All") %>%
    filter(cat_soc_care_recipient!="Both") %>%
    ggplot(aes(x=val_cohort_year,y=amt_spells_mean,ymin=amt_spells_ci_lo, ymax=amt_spells_ci_hi, group=cat_soc_care_recipient, shape=cat_soc_care_recipient, colour=cat_soc_care_recipient)) +
  geom_point(size=rel(2)) +
  geom_line() +
  geom_errorbar(width = 0.2) +
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_colour_manual(values = order_colours_adr_s[1:2]) +
  labs(x=NULL,y=NULL,title="Mean hospital spells")
)
```

### Total days in hospital

```{r}
(subplot_hospital_stay <-
    los_by_recipient %>%
    filter(cat_time==365) %>%
    filter(val_cohort_year!="All") %>%
    filter(cat_soc_care_recipient!="Both") %>%
    ggplot(aes(x=val_cohort_year, y=amt_los_mean,ymin=amt_los_ci_lo, ymax=amt_los_ci_hi, group=cat_soc_care_recipient, shape=cat_soc_care_recipient, colour=cat_soc_care_recipient)) +
  geom_point(size=rel(2)) +
  geom_line() +
  geom_errorbar(width = 0.2) +
  theme(plot.title = element_text(hjust = 0.5)) +
 scale_colour_manual(values = order_colours_adr_s[1:2]) +
  labs(x=NULL,y=NULL,title="Mean hospital spells")
)
```

### Proportion admitted

```{r}
(subplot_hospital_admitted <-
  percent_admitted_by_recipient %>%
  filter(cat_time==365) %>%
  filter(val_cohort_year!="All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  filter(cat_dataset=="smr01") %>%
  ggplot(aes(x=val_cohort_year, y=perc_admitted, group=cat_soc_care_recipient, shape=cat_soc_care_recipient, fill=cat_soc_care_recipient)) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0, 1), n.breaks = 9) +
  theme(plot.title = element_text(hjust = 0.5)) +
   scale_fill_manual(values = order_colours_adr_s[1:2]) +
  labs(x=NULL,y=NULL,title="% with inpatient admissions\nin final 12 months of life")
)
```

### ICD chapters

```{r}
(subplot_icd_chapters <-
  icd_chapters_by_recipient %>%
  filter(val_cohort_year!="All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  ggplot(aes(x=val_cohort_year, y=perc, group=cat_soc_care_recipient, shape=cat_soc_care_recipient, fill=cat_soc_care_recipient)) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
  scale_y_continuous(labels = scales::label_percent(), limits = c(0, 1), n.breaks = 9) +
  theme(plot.title = element_text(hjust = 0.5)) +
   scale_fill_manual(values = order_colours_adr_s[1:2]) +
  labs(x=NULL,y=NULL,title="% with inpatient admissions\nin final 12 months of life")
)
```



# Save figures

- The default width of a columnar section is 38.36cm
- So a plot width of 18cm leaves space for a gap!

```{r}
save_ggplot <- function(plot, filename, ...) {
  ggsave(filename = file.path(dir_output, filename), plot = plot, bg = "white", dpi = 600, ...)
}

tribble(
  ~plot, ~filename,
  subplot_age, "fig_age.png",
  subplot_sex, "fig_sex.png",
  subplot_comorbidities, "fig_comorbidities.png",
  subplot_hospital_admitted, "fig_comorbidities.png"
) %>%
  pwalk(.f = function(plot, filename) save_ggplot(plot = plot, filename = filename, units = "cm", width = 18, height = 12))

```

