---
title: "Social care exploration scripts"
author: "Jan Savinc"
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

# Preliminaries

## Packages

```{r}
library(tidyverse)
library(patchwork)
library(broom)  # for tidy()
library(tictoc)  # for timing
library(janitor)  # for some nice cleaning functions

source("./FUNCTIONS.R")  # load utility functions
```

## Functions & constants

```{r}
# where to save outputs
dir_output <- "C:/Users/40011625/OneDrive - Edinburgh Napier University/SCADR/covid_mortality_home/drafts/ADR UK 2025/"

# specify the order of place of death
levels_place_of_death <- c("Hospital", "Home", "Care home", "All")

code_place_of_death_consistently <- function(data_tbl) {
  data_tbl %>%
    mutate(
      ## shorten to "Home"
      cat_place_of_death = if_else(cat_place_of_death == "Home & non-institution", true = "Home", false = cat_place_of_death),
      ## Shorten to "Care home"
      cat_place_of_death = if_else(cat_place_of_death == "Care home & other", true = "Care home", false = cat_place_of_death),
      cat_place_of_death = factor(cat_place_of_death, levels = levels_place_of_death)
    )
}
```

## Load data

```{r}
## load cohort size by PoD
cohort_by_pod <- 
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/outputs/cohort_size_pod.csv") %>%
  repeat_data_adding_a_catchall_category(var_to_change = cat_place_of_death, label_for_catchall = "All") %>%
  group_by(val_cohort_year, cat_place_of_death) %>%
  summarise(n=sum(n), .groups = "drop")

crosstab_record_type <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/crosstab_individuals_by_type_of_source_record.csv")  # rounded to nearest 10!

crosstab_by_year <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/crosstab_records_individuals_source_files_by_financial_year_quarter.csv")

tbl_proportion_any_soc_care_record <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/tbl_count_prop_pod_year_any_source_record.csv") %>%   # rounded to nearest 10
  code_place_of_death_consistently
  
tbl_proportion_home_care_record <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/tbl_count_prop_pod_year_home_care_record.csv") %>%
  code_place_of_death_consistently

tbl_proportion_client_characteristics <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/tbl_proportions_decedents_client_characteristics.csv") %>%
  code_place_of_death_consistently

descriptives_number_home_care_records <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/freq_home_care_records_decedents.csv") %>%
  code_place_of_death_consistently

descriptives_mean_number_home_care_records <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/descriptives_home_care_records_decedents.csv")

descriptives_mean_duration_home_care_records <- read_csv("X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_exploration/descriptives_duration_home_care_records_decedents.csv") %>%
  code_place_of_death_consistently
```

# Outputs

## Numbers of records, total

Note 2025-09-01: this table is correct in reporting the number of unique individuals for the entire dataset; the below table of records over time gives the potentially misleading impression that there are more records over time, but some of those will be repeated for the same individuals.

```{r}
tbl_number_of_records <-
  crosstab_record_type %>%
  summarise(
    num_client = sum(n[bin_client]),
    num_demographics = sum(n[bin_demographics]),
    num_homecare = sum(n[bin_homecare]),
    num_iorn = sum(n[bin_iorn])
  )
```

## Numbers of records, over time

Note that because this counts unique individuals by financial quarter, it may appear to overcount the number of individuals in the study.

The 2017-18 discrepancy between the number of individuals/records in the client file vs demographics/homecare is likely an issue with data collection at the very start of SOURCE data collection;
The gradual decline in the number of records from 2019-20 to 2020-21 is due to people in the cohort dying as the cohort was defined based on death date with a cut-off at 2021-03-22.
In other words, the figure in 2019-4, for example, is a mix of people who died in 2019-4, or in the next 4 quarters; in 2020-21, the figure only captures individuals who would die in the next 3 quarters, and so on.

```{r}
(fig_records_over_time <-
  crosstab_by_year %>%
  filter(dt_financial_year!="all") %>%
  mutate(across(c(dt_financial_year, dt_financial_quarter), as.integer)) %>%
  pivot_longer(cols = -c(dt_financial_year, dt_financial_quarter), names_to = c("measure", "dataset"), names_pattern = "(n\\_records|n\\_individuals)\\_(client|demographics|homecare|iorn)", values_to = "n") %>%
  mutate(
    dt_str = str_c(dt_financial_year, dt_financial_quarter, sep = "-"),
    measure = if_else(measure=="n_records", "Records", "Individuals")
    ) %>%
  ggplot(
    aes(x=dt_str, y=n, colour=dataset, shape=dataset)
  ) +
    facet_grid(vars(measure)) +
    geom_point(size = rel(2)) +
    geom_line(aes(group = dataset)) +
    scale_colour_viridis_d(end = 0.8) +
   scale_y_continuous(labels = scales::label_comma()) +
   theme(legend.position = "top") +
   labs(
     x = NULL, y = NULL
   )
 )
```

## Number of individuals over time

```{r}
(fig_individuals_over_time <-
  crosstab_by_year %>%
  filter(dt_financial_year!="all") %>%
  mutate(across(c(dt_financial_year, dt_financial_quarter), as.integer)) %>%
  pivot_longer(cols = -c(dt_financial_year, dt_financial_quarter), names_to = c("measure", "dataset"), names_pattern = "(n\\_records|n\\_individuals)\\_(client|demographics|homecare|iorn)", values_to = "n") %>%
  mutate(
    dt_str = str_c(dt_financial_year, dt_financial_quarter, sep = "-"),
    measure = if_else(measure=="n_records", "Records", "Individuals")
    ) %>%
   filter(measure == "Individuals") %>%
  ggplot(
    aes(x=dt_str, y=n, colour=dataset, shape=dataset)
  ) +
    geom_point(size = rel(2)) +
    geom_line(aes(group = dataset)) +
    scale_colour_viridis_d(end = 0.8) +
   scale_y_continuous(labels = scales::label_comma()) +
   theme(legend.position = "top") +
   labs(
     x = NULL, y = NULL
   )
 )

ggsave(fig_individuals_over_time, filename = str_c(dir_output, "/fig_individuals_over_time_by_dataset.png"), bg = "white", dpi = 300, width = 7, height = 4)
```


## Proportion of people with home care records, by time

```{r}
fig_num_homecare <-
  tbl_proportion_home_care_record %>%
  filter(cat_place_of_death=="All") %>%
  ggplot(aes(x = val_cohort_year, y = n)) +
  geom_point(size = rel(2)) +
  geom_line(aes(group=cat_place_of_death)) +
  scale_y_continuous(label = scales::label_comma(), breaks = scales::breaks_pretty(n=5), limits = c(0,NA)) +
  labs(y = NULL, x = NULL, title = "Number of home care clients") +
  NULL

fig_prop_homecare <-
  tbl_proportion_home_care_record %>%
  filter(cat_place_of_death=="All") %>%
  ggplot(aes(x = val_cohort_year, y = prop)) +
  geom_point(size = rel(2)) +
  geom_line(aes(group=cat_place_of_death)) +
  scale_y_continuous(label = scales::label_percent(), breaks = scales::breaks_pretty(n=5), limits = c(0,NA)) +
  labs(y = NULL, x = NULL, title = "Percent of home care clients") +
  NULL

fig_homecare <-
  wrap_plots(
    fig_num_homecare, fig_prop_homecare,
    guides = "collect"
  )

ggsave(fig_homecare, filename = str_c(dir_output,"/fig_homecare.png"), bg = "white", dpi = 300, width = 7, height = 4)
```


## Proportion of people with home care records, by PoD & time

```{r}
annotations_homecare_by_pod <-
  tbl_proportion_home_care_record %>%
  filter(cat_place_of_death!="All") %>%
  filter(val_cohort_year == "2018-19") %>%
  mutate(y = if_else(cat_place_of_death == "Home", n+700, n-700))

fig_prop_homecare_by_pod <-
  tbl_proportion_home_care_record %>%
  filter(cat_place_of_death!="All") %>%
  ggplot(aes(x = val_cohort_year, y = prop, colour = cat_place_of_death, shape = cat_place_of_death)) +
  geom_point(size = rel(2)) +
  geom_line(aes(group = cat_place_of_death)) +
  scale_y_continuous(label = scales::label_percent(), limits = c(0, NA), breaks = scales::breaks_pretty(n=5)) +
  labs(y = NULL, x = NULL, title = "Percent of home care clients", subtitle = "Amongst all decedents") +
  scale_colour_viridis_d(end = 0.8) +
  NULL

fig_num_homecare_by_pod <-
  tbl_proportion_home_care_record %>%
  filter(cat_place_of_death!="All") %>%
  ggplot(aes(x = val_cohort_year, y = n, colour = cat_place_of_death, shape = cat_place_of_death)) +
  geom_point(size = rel(2)) +
  geom_line(aes(group = cat_place_of_death)) +
  scale_y_continuous(label = scales::label_comma(), breaks = scales::breaks_pretty(n=5), limits = c(0, NA)) +
  geom_text(
    data = annotations_homecare_by_pod,
    aes(x=val_cohort_year,y=y,label=cat_place_of_death,colour=cat_place_of_death), hjust = 0, angle = 15
  ) +
  labs(y = NULL, x = NULL, title = "Number of home care clients") +
  scale_colour_viridis_d(end = 0.8) +
  NULL

fig_homecare_by_pod <-
  wrap_plots(
    fig_num_homecare_by_pod, fig_prop_homecare_by_pod, guides = "collect"
  ) & theme(legend.position = "none")

ggsave(fig_homecare_by_pod, filename = str_c(dir_output,"/fig_homecare_by_pod.png"), bg = "white", dpi = 300, width = 7, height = 4)
```


## Proportion of place of death, by home care

```{r}
## compute proportion of PoD by whether people received home care or not
tbl_proportion_pod_by_homecare <-
  tbl_proportion_home_care_record %>%
  filter(cat_place_of_death!="All") %>%  # doesn't make sense to have All as place of death when computing distribution!
  mutate(n_no_homecare = denominator - n) %>%
  rename(true = n, false = n_no_homecare) %>%
  # select(-c(prop, denominator)) %>%
  pivot_longer(cols = c(true, false), names_to = "bin_received_homecare", values_to = "n", names_transform = as.logical) %>%
  group_by(val_cohort_year, bin_received_homecare) %>%
  mutate(denominator = sum(n), prop = n/denominator) %>%
  ungroup

annotations_pod_by_homecare <-
  tbl_proportion_pod_by_homecare %>%
  # filter(val_cohort_year=="2018-19") %>%
  mutate(
    y = case_when(
      # complicated-looking logic to put text in right place relative to points & lines
      (cat_place_of_death == "Hospital" & bin_received_homecare) | (cat_place_of_death != "Hospital" & !bin_received_homecare) ~ prop + 0.05,
      (cat_place_of_death == "Hospital" & !bin_received_homecare) | (cat_place_of_death != "Hospital" & bin_received_homecare) ~ prop - 0.05,
      TRUE ~ NA_real_
    ),
    label = case_when(
      cat_place_of_death == "Hospital" & bin_received_homecare & val_cohort_year=="2018-19" ~ "Home care",
      cat_place_of_death == "Hospital" & !bin_received_homecare & val_cohort_year=="2018-19" ~ "No home care",
      val_cohort_year!="2018-19" ~ scales::label_percent(accuracy=1)(prop),
      TRUE ~ ""
    ),
  )

(fig_proportion_pod_by_homecare <-
  tbl_proportion_pod_by_homecare %>%
  ggplot(data = ., aes(x = val_cohort_year, y = prop, group = bin_received_homecare, colour = bin_received_homecare, shape = bin_received_homecare)) +
    geom_point(size=rel(2)) +
    geom_line() +
    facet_wrap(vars(cat_place_of_death)) +
    scale_y_continuous(label = scales::label_percent(), limits = c(0, NA), breaks = scales::breaks_pretty(n=5)) +
    geom_text(data = annotations_pod_by_homecare, aes(x=val_cohort_year,y=y,label=label,colour=bin_received_homecare), hjust = 0, angle = -17) +
    labs(y = NULL, x = NULL) +
    theme(legend.position = "none") +
    NULL
)

ggsave(
  fig_proportion_pod_by_homecare, bg = "white", dpi = 300, filename = str_c(dir_output, "/fig_pod_by_homecare_receipt.png"), width = 7, height = 4
)
```

## Client characteristics over time

```{r}
fig_client_characteristics <-
  tbl_proportion_client_characteristics %>%
  select(1:2, matches("prop$")) %>%
  pivot_longer(cols = matches("bin\\_"), names_to = "characteristic", values_to = "prop") %>%
  mutate(
    characteristic = case_when(
      str_detect(characteristic, "pall") ~ "Received palliative care",
      str_detect(characteristic, "alone") ~ "Lived alone",
      str_detect(characteristic, "unpaid") ~ "Received unpaid care",
      TRUE ~ NA_character_
    )
  ) %>%
  ggplot(aes(x = val_cohort_year, y = prop, colour = characteristic, shape = characteristic)) +
  facet_wrap(vars(cat_place_of_death), nrow = 1) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 0.1), limits = c(0,0.1), breaks = scales::breaks_pretty(n=6)) +
  geom_point(size=rel(2)) +
  geom_line(aes(group = characteristic)) +
  labs(y = NULL, x = NULL, title = "Client characteristics") +
  scale_colour_viridis_d(end = 0.8) +
  NULL

ggsave(fig_client_characteristics, filename=str_c(dir_output,"/fig_client_characteristics.png"), bg="white", dpi = 300, width = 11, height = 5)
```

### Client characteristics for all of Scotland

```{r}
data_client_characteristics <-
  tbl_proportion_client_characteristics %>%
  select(1:2, matches("prop$")) %>%
  pivot_longer(cols = matches("bin\\_"), names_to = "characteristic", values_to = "prop") %>%
  mutate(
    characteristic = case_when(
      str_detect(characteristic, "pall") ~ "Received palliative care",
      str_detect(characteristic, "alone") ~ "Lived alone",
      str_detect(characteristic, "unpaid") ~ "Received unpaid care",
      TRUE ~ NA_character_
    )
  )

annotations_client_characteristics_scotland <-
  data_client_characteristics %>%
  filter(val_cohort_year == "2018-19" & cat_place_of_death == "All") %>%
  mutate(
    # y = if_else(characteristic == "Received palliative care", prop+0.005, prop+0.007),
    y = prop+0.007,
    angle = if_else(characteristic == "Received palliative care", -5, 10)
  )
  
(fig_client_characteristics_scotland <-
  data_client_characteristics %>%
  filter(cat_place_of_death=="All") %>%
  ggplot(aes(x = val_cohort_year, y = prop, colour = characteristic, shape = characteristic)) +
  # scale_y_continuous(labels = scales::label_percent(accuracy = 0.1), limits = c(0,0.1), breaks = scales::breaks_pretty(n=6)) +
  scale_y_continuous(labels = scales::label_percent(accuracy = 1), limits = c(NA,0.1), breaks = scales::breaks_pretty(n=6)) +
  geom_point(size=rel(2)) +
  geom_line(aes(group = characteristic)) +
  geom_text(data = annotations_client_characteristics_scotland, aes(x=val_cohort_year, y=y, label=characteristic, angle = angle), hjust = 0) +
  labs(y = NULL, x = NULL, title = "Client characteristics") +
  scale_colour_viridis_d(end = 0.8) +
  theme(legend.position = "none") +
  NULL)

ggsave(fig_client_characteristics_scotland, filename=str_c(dir_output,"/fig_client_characteristics_scotland.png"), bg="white", dpi = 300, width = 7, height = 4)
```

