---
title: "Social Care analysis script for IPDLN 2026 poster"
author: "Jan Savinc"
format: html
editor: source
editor_options: 
  chunk_output_type: console
---

# Preliminaries

## Packages

```{r}
library(tidyverse)
library(patchwork)
library(broom)  # for tidy()
library(tictoc)  # for timing
library(janitor)  # for some nice cleaning functions

source("./FUNCTIONS.R")  # load utility functions
```

## Functions & constants

```{r}
# where to save outputs
dir_output <- "C:/Users/40011625/OneDrive - Edinburgh Napier University/SCADR/covid_mortality_home/dissemination/drafts/IPDLN 2026/"

# specify the order of place of death
levels_place_of_death <- c("Hospital", "Home", "Care home", "All")

code_place_of_death_consistently <- function(data_tbl) {
  data_tbl %>%
    mutate(
      ## shorten to "Home"
      cat_place_of_death = if_else(cat_place_of_death == "Home & non-institution", true = "Home", false = cat_place_of_death),
      ## Shorten to "Care home"
      cat_place_of_death = if_else(cat_place_of_death == "Care home & other", true = "Care home", false = cat_place_of_death),
      cat_place_of_death = factor(cat_place_of_death, levels = levels_place_of_death)
    )
}
```

## Load data

```{r}
demographics_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_characteristics_of_soc_care_recipients_vs_nonrecipients.csv") %>%
  mutate(val_cohort_year = factor(val_cohort_year, levels = order_cohort_year))

comorbidities_by_recipient <-
  read_csv(file = "X:/R2090/2021-0312 Deaths at home/safe_haven_exports/social_care_analysis/tbl_freq_comorbidities_recipients_vs_nonrecipients.csv") %>%
  mutate(val_cohort_year = factor(val_cohort_year, levels = order_cohort_year))

# list the variables in the demographics table
demographics_by_recipient %>% count(variable)
```

## Wrangle data

### Consistent place of death name in demographics

```{r}
demographics_by_recipient <-
  demographics_by_recipient %>%
  mutate(
    level = case_match(level,
                       "Home & non-institution" ~ "Home",
                       "Care home & other" ~ "Care home",
                       .default = level)
  )
```


### Compute year-on-year changes in demographics

Merge the previous years' proportion and compute the difference. Start by deriving the previous year!

```{r}
demographics_by_recipient <-
  demographics_by_recipient %>%
  mutate(prev_year = order_cohort_year[as.integer(val_cohort_year)-1L]) %>%
  left_join(demographics_by_recipient %>% select(val_cohort_year:level,prop_previous = prop), by = join_by(prev_year==val_cohort_year, cat_soc_care_recipient, variable, level)) %>%
  mutate(
    prop_previous = if_else(val_cohort_year=="All", NA_real_, prop_previous),  # set to NA for the All years to avoid nonsensical results
    prop_change = prop - prop_previous
    ) %>%
  select(-c(prev_year,prop_previous))
```


### Compute year-on-year changes in comorbidities

Merge the previous years' proportion and compute the difference. Start by deriving the previous year!

```{r}
comorbidities_by_recipient <-
  comorbidities_by_recipient %>%
  mutate(prev_year = order_cohort_year[as.integer(val_cohort_year)-1L]) %>%
  left_join(comorbidities_by_recipient %>% select(val_cohort_year:cat_comorbidity,perc_previous = perc), by = join_by(prev_year==val_cohort_year, cat_soc_care_recipient, cat_comorbidity)) %>%
  mutate(
    perc_previous = if_else(val_cohort_year=="All", NA_real_, perc_previous),  # set to NA for the All years to avoid nonsensical results
    perc_change = perc - perc_previous
    ) %>%
  select(-c(prev_year,perc_previous))
```



### Top 12 most frequent comorbidities

I derive these from the "All" cohort year and "Both" social care recipient categories...

```{r}
top_12_comorb <-
  comorbidities_by_recipient %>%
  filter(val_cohort_year=="All") %>%
  count(cat_comorbidity, wt = n) %>%
  slice_max(order_by = n, n = 12) %>% pull(cat_comorbidity)
```


# Outputs

## Age by recipients

- ethnicity doesn't make for a good graph...
- 

```{r}
# TODO: arrange multiple of these in a grid using patchwork
demographics_by_recipient %>% 
  filter(variable == "cat_age_decades") %>%
  filter(val_cohort_year == "All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  ggplot(
    aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
  ) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10")

# ethnicity not very visualisable
demographics_by_recipient %>% 
  filter(variable == "cat_ethnic_group_collapsed") %>%
  filter(val_cohort_year == "All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  filter(level!="White") %>%
  ggplot(
    aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
  ) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10")

demographics_by_recipient %>% 
  filter(variable == "cat_marital_status_condensed") %>%
  filter(val_cohort_year == "All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  ggplot(
    aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
  ) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10")

demographics_by_recipient %>% 
  filter(variable == "cat_sex") %>%
  filter(val_cohort_year == "All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  ggplot(
    aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
  ) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10")

demographics_by_recipient %>% 
  filter(variable == "cat_simd_quintile") %>%
  filter(val_cohort_year == "All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  ggplot(
    aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
  ) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10")

demographics_by_recipient %>% 
  filter(variable == "cat_ur2") %>%
  filter(val_cohort_year == "All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  ggplot(
    aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
  ) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10")

demographics_by_recipient %>% 
  filter(variable == "cat_place_of_death") %>%
  filter(val_cohort_year == "All") %>%
  filter(cat_soc_care_recipient!="Both") %>%
  ggplot(
    aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
  ) +
  geom_col(position = position_dodge(width = 0.9), colour = "grey10")
```

## Demographics over time

- in 2020/21 age distribution has changed very little
- in 2020/21 sex shifted towards higher proportion in males in both recipients and non-recipients
- in 2020/21 SIMD has seen a decrease in 2nd quintile and increase in 1st quintile, while 3-5 remained similar
- not much changed on marital status


```{r}
subplot_data_extract <- function(which_var) {
  demographics_by_recipient %>% 
      filter(variable == which_var) %>%
      filter(val_cohort_year == "2020-21") %>%
      filter(cat_soc_care_recipient!="Both") %>%
      mutate(level = factor(level, levels=unique(level)))  # set order of factor levels as it was exported
}

subplot_ggplot <- function(data, y = "prop") {
  ggplot(
    data = data,
    aes(x = level, y = .data[[y]], fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
    ) +
    geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
    scale_y_continuous(labels = scales::label_percent()) +
    labs(x=NULL,y=NULL) +
    theme(plot.title = element_text(hjust = 0.5))
}

(subplot_age <-
  wrap_plots(
    subplot_data_extract("cat_age_decades") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_age_decades") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Age groups")

(subplot_sex <-
  wrap_plots(
    subplot_data_extract("cat_sex") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_sex") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Sex")

(subplot_marstat <-
  wrap_plots(
    subplot_data_extract("cat_marital_status_condensed") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_marital_status_condensed") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Marital status")

(subplot_simd <-
  wrap_plots(
    subplot_data_extract("cat_simd_quintile") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_simd_quintile") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Deprivation (SIMD quintile)")

(subplot_ur2 <-
  wrap_plots(
    subplot_data_extract("cat_ur2") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_ur2") %>%
      subplot_ggplot(y = "prop_change") +
      scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Urban-rural location")

(subplot_pod <-
  wrap_plots(
    subplot_data_extract("cat_place_of_death") %>%
      subplot_ggplot() +
      labs(title="2020-21")
    ,
    subplot_data_extract("cat_place_of_death") %>%
      subplot_ggplot(y = "prop_change") +
      # scale_y_continuous(labels = scales::label_percent(), limits = c(-0.02, 0.02)) +
      labs(title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    )) &
    plot_annotation(title = "Place of death")

# TODO: I lose the plot_annotation title for the subplots in this step...
wrap_plots(
  list(
    subplot_age, subplot_sex,
    subplot_pod, subplot_marstat,
    subplot_simd, subplot_ur2
  ), ncol = 2, nrow = 3
)

# TODO: below is cruft, remove

# demographics_by_recipient %>% 
#   filter(variable == "cat_age_decades") %>%
#   filter(val_cohort_year == "2020-21") %>%
#   filter(cat_soc_care_recipient!="Both") %>%
#   pivot_longer(cols = c(prop, prop_change), names_to = "period", values_to = "prop") %>%
#   mutate(period = case_match(period, "prop"~"2020-21", "prop_change"~"change from 2019-20")) %>%
#   ggplot(
#     aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
#   ) +
#   geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
#   facet_wrap(vars(period, val_cohort_year), scales = "free_y")
# 
# 
# demographics_by_recipient %>% 
#   filter(variable == "cat_marital_status_condensed") %>%
#   filter(val_cohort_year != "All") %>%
#   filter(cat_soc_care_recipient!="Both") %>%
#   ggplot(
#     aes(x = level, y = prop, fill = cat_soc_care_recipient, group = cat_soc_care_recipient)
#   ) +
#   geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
#   facet_wrap(vars(val_cohort_year))
```

## Comorbid conditions over time

Only a few comorbidities that were more common in social care non-recipients... the individuals affected likely die before reaching social care age.

- solid tumour without metastasis
- metastatic cancer
- alcohol abuse
- liver disease
- lymphoma
- psychoses
- drug abuse
- aids/hiv

```{r}
# prints the comorbidities more common in non-recipients
comorbidities_by_recipient %>% select(-c(n, perc_change)) %>% pivot_wider(names_from = cat_soc_care_recipient, values_from = perc) %>% filter(Recipient < `Non-recipient`) %>% arrange(desc(Recipient)) %>% print_all

# TODO: decide how to present this... only the most frequent comorbs, but also the ones with the biggest difference between recipients and non-recipients?
# TODO: changes in pandemic weren't massive, not sure how to present those...

(subplot_comorbidities <-
  wrap_plots(
    comorbidities_by_recipient %>% 
      filter(val_cohort_year == "2020-21") %>%
      filter(cat_comorbidity %in% top_12_comorb) %>%
      mutate(cat_comorbidity = factor(cat_comorbidity, levels=unique(cat_comorbidity))) %>%  # set order of factor levels as it was exported
      ggplot(aes(x=cat_comorbidity,y=perc,fill=cat_soc_care_recipient,group=cat_soc_care_recipient)) +
      geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
      scale_y_continuous(labels = scales::label_percent()) +
      theme(plot.title = element_text(hjust = 0.5)) +
      labs(x=NULL,y=NULL,title="2020-21")
    ,
    comorbidities_by_recipient %>% 
      filter(val_cohort_year == "2020-21") %>%
      filter(cat_comorbidity %in% top_12_comorb) %>%
      mutate(cat_comorbidity = factor(cat_comorbidity, levels=unique(cat_comorbidity))) %>%  # set order of factor levels as it was exported
      ggplot(aes(x=cat_comorbidity,y=perc_change,fill=cat_soc_care_recipient,group=cat_soc_care_recipient)) +
      geom_col(position = position_dodge(width = 0.9), colour = "grey10") +
      scale_y_continuous(labels = scales::label_percent()) +
      theme(plot.title = element_text(hjust = 0.5)) +
      labs(x=NULL,y=NULL,title="Percentage point change\nfrom 2019-20")
    ,
    guides = "collect"
    ) &
    plot_annotation(title = "Comorbid conditions")
)
```

